% !TEX root = ../thesis-example.tex
%
\pdfbookmark[0]{Acknowledgments}{Acknowledgments}
\chapter*{Acknowledgments}
\label{sec:thanks}
\vspace*{-10mm}


First of all, I am extremely grateful to my thesis supervisors at {\thesisUniversity}, \textbf{Prof. Dr. Bernd Möbius} and \textbf{Dr. Jürgen Trouvain}; this thesis could not have been written without their support, encouragement, and feedback.

The work reported in this thesis was partially supported by the Franco-German project \textit{Individualized Feedback for Computer-Assisted Spoken Language learning} (IFCASL), funded jointly by the Deutsche Forschungsgemeinschaft (DFG) and the Agence Nationale de la Recherche (ANR). I thank the the entire \textbf{Project IFCASL team} at both Saarland University (Saarbrücken, Germany) and LORIA (Nancy, France) for giving me the opportunity to work with them on this project over the last two years, and for supporting my work at both institutions.

Among the researchers at LORIA, I would especially like to thank:
\textbf{Dr. Yves Laprie} for providing invaluable information on the automatic processing of prosody and for generously sharing his office with me;
\textbf{Dr. Julie Busset} for allowing me to work closely with her on refactoring the JSnoori software; % to facilitate integration with programs such as de-stress};
\textbf{Dr. Anne Bonneau} for advising me on feedback on prosody in nonnative speech and for welcoming me in Nancy;
%prosody feedback in Computer-Assisted Language Learning (CAPT);
and 
\textbf{Dr. Dominique Fohr} and \textbf{Dr. Odile Mella} for elucidating the automatic segmentation of nonnative speech. 
%Dr. Julie Busset, Dr. Vincent Colotte,  and Dr. Yves Laprie 
%at LORIA, who provided invaluable advice on prosody in Computer-Assisted Language Learning (CAPT), speech processing, and integration of the \textit{JSnoori} software with the \textit{de-stress} system. 

I am also extremely grateful to IFCASL team members \textbf{Jeanin Jügler} and \textbf{Dr. Frank Zimmerer} at {\thesisUniversity} for all their help, and especially for assisting with the lexical stress error annotation project and providing the German translation of the thesis abstract. 

I would additionally like to thank the \textbf{annotators} (not named to preserve anonymity) who donated their time to the error annotation project; without the labeled data produced through their efforts, much of the work reported in this thesis would not have been possible.

My heartfelt thanks also go to \textbf{Marie Springinsfeld} for generously providing the French translation of the thesis abstract, and to \textbf{Nathaneil Ward} and \textbf{Max Rabkin} for kindly helping to proofread this document and providing valuable feedback.

Finally, I would like to express my gratitude to 
my mother, \textbf{Eileen Julian}, and to {my wonderful friends} in Saarbrücken and around the world.
%my friends and family, 
%especially Eileen Julian, James Newman, Max Rabkin, Rebecca Row, and Nat Ward, 
I could not have completed this thesis or the studies leading up to it without 
their loving support.